---
title: "Destroy `observeEvent`s"
author: "Jonas Hagenberg"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Rationale
- `observeEvent`s stay registered and are duplicated when a module with the
same `id` is added again
- many registered `observeEvent`s can negatively impact the performance

# Solution
- use the `destroy` method of an `observeEvent`
- in order to access the `destroy` method of an `observeEvent`,
store the `observeEvent` to `session$userData`
- check out [Appsilon's blog post](https://appsilon.com/how-to-safely-remove-a-dynamic-shiny-module/) on this topic

## Utility function
If you follow the naming convention of `{module id}_observer_{observe name}` for the `observeEvent`s, you can use the following function to destroy the `observeEvent`s when you delete a module. `id` is the module `id` and for `.session` you pass the `session` object.

Please look at `example_destroy_observeevent.R` for a working example.

```{r, eval = FALSE}
remove_observers <- function(id, .session) {
  invisible(
    lapply(grep(paste0(id, "_observer"), names(.session$userData), value = TRUE),
           function(i) {
             .subset2(.session$userData, i)$destroy()
           })
  )
}
```