---
title: "Dynamically adding and removing Shiny modules"
author: "Jonas Hagenberg"
institute: "Appsilon Shiny Conference"
date: "27.04.2022"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#1a5f96",
  secondary_color = "#03A696",
  code_inline_background_color = "#f8f8f8"
)
```

# Recap: Shiny modules
- reusable pieces of Shiny code
- a module consists of an UI and a server part
- can think of it as a "mini Shiny app"
- can be used several times within an app
- every module instance needs an `id` to be unique within the app

---
# Module structure - main app
```{r, eval = FALSE}
# define the UI part
ui <- fluidPage(
  # other UI elements
  # ...
  # module UI
* module_UI(id = "id_1")
)

# define the server part
server <- function(input, output,
                   session) {
  # other logic
  # ...
  # module logic
* module_server(id = "id_1")
}

shinyApp(ui, server)
```

???
- you call the module UI part (here called `module_UI`) in the main app UI function
- you call the module server part (here called `module_server`) in the server
function of the main app
- as you can see, they need an ID to make them unique and connect the server & UI part

---
# Module structure - modules
```{r, eval = FALSE}
# define the UI part
module_UI <- function(id) {
* ns <- NS(id)
  
  tagList(
    # your UI elements here
  )
}

# define the server part
module_server <- function(id) {
* moduleServer(
    id,
    function(input, output, session) {
      # your logic here
    }
  )
}
```
???
- scaffold of a module
- in the UI part you need take care of the namespacing so that everything is
unique for this instance of the module
- the server part looks very similar to main app server function

---
# Dynamic modules
- modules are either called when writing the Shiny app or dynamically
--

- dynamic modules can be useful when:
  - something has to be repeated based on user input
  - e.g. show several plots or tables
  - a set of UI elements
--

- can become complex easily -> use it with care!
  
???
- set of UI elements: e.g. additional layers in plots that can be controlled

---
class: inverse, center, middle

### Livecoding: dynamically add a module

???
- how to get from static modules to dynamically added modules?

---
# Adding modules
When adding a module:
- call the module `server` function
- insert the module UI elements with `insertUI`
- use an appropriate `selector` where to insert the UI - one can use an empty `div`
with an `id`
- use a different `id` for every added module instance

---
# How to remove a module
- remove UI part with `removeUI`
- needs to provide a selector
- works well for one UI element but not for several in module UI
- wrap UI elements with `div` with module `id` instead of `tagList`

```{r, eval = FALSE}
module_UI <- function(id) {
  ns <- NS(id)
  div( #<<
    id = id,
    # your UI elements here
  )
}
```

???
- the selector is based on the id of the UI elements
- in our case the id of the module

---
class: inverse, center, middle

### Livecoding: dynamically remove a module

---
# Recap: remove UI of a module
- use `removeUI`
- create a selector with `#` and the module `id`
- use a naming scheme for the modules that can be easily incremented/decremented

???
- this removes the contents of the module on the client side

---
# How to remove the server part
- so far, only UI elements were removed
- `input`s, `reactive`s and `observeEvent`s of the module are still in
the server part
- can negatively influence the performance
- therefore also remove these

---
# Remove `input`s

???
- `input`s would be recycled if a module with the same name is created again
- currently don't know a way to remove `reactive` expressions

--

```{r, eval = FALSE}
remove_shiny_inputs <- function(id, .input) {
  invisible(
    lapply(grep(id, names(.input), value = TRUE), function(i) {
      .subset2(.input, "impl")$.values$remove(i) #<<
    })
  )
}
```

By [Thomas Roh](https://roh.engineering/posts/2020/02/shiny-add/removing-modules-dynamically/), CC BY-NC 4.0

???
- pass the `id` of the module that should be removed (`id`), the different inputs
are matched automatically
- pass the list (here: `input`) where the inputs are stored
- for every module input, call the `remove` method
- rather hacky method, not Shiny official

---
class: inverse, center, middle

### Livecoding: remove inputs